---
title: "Day-11-12"
author: Caleb Cripe
format: html
editor: visual
---

### **Exploratory Data Analysis and Linear Regression in R**

In this assignment, you will again analyze the `airquality` dataset using various statistical tests, data transformation techniques, and regression modeling. Follow the step-by-step guiding questions to complete the analysis in a qmd.

## **Part 1: Normality Testing**

1.  Load the airquality dataset in R. What does this dataset represent? Explore its structure using functions like `str()` and `summary()`.

```{r}
data(airquality)
str(airquality)
summary(airquality)
```

This dataset represent the records for four variables that impact daily air quality in New York City. The four variables it accounts for are Ozone, Solar.R, Wind, and Temp.

2.  Perform a Shapiro-Wilk normality test on the following variables: `Ozone`, `Temp`, `Solar.R`, and `Wind.`

```{r}
shapiro.test(airquality$Ozone)

shapiro.test(airquality$Solar.R)

shapiro.test(airquality$Temp)

shapiro.test(airquality$Wind)
```

3.  What is the purpose of the Shapiro-Wilk test?

The Shapiro-Wilk test is intended to test data for normal distribution in relation to the null hypothesis, with low p-values indicating that the data is likely not normally distributed. 

4.  What are the null and alternative hypotheses for this test?

The null hypothesis for the Shapiro-Wilk test is that the data is normally distributed, with a p-value < 0.05 rejecting this, and the alternative hypothesis is that the data is not normally distributed. 

5.  Interpret the p-values. Are these variables normally distributed?

Out of the four variables only wind appears to be not normally distributed according to the Shapiro-Wilk test, while Ozone, Solar.R, and Wind are normally distributed. 

## **Part 2: Data Transformation and Feature Engineering**

6.  Create a new column with `case_when` traslating the Months into four seasons (Winter (Nov, Dec, Jan), Spring (Feb, Mar, Apr), Summer (May, Jun, Jul), and Fall (Aug, Sep, Oct)).

```{r}
library(dplyr)

airquality <- airquality %>%
  mutate(season = case_when(
    Month %in% c(2, 3, 4) ~ "Spring",
    Month %in% c(5, 6, 7) ~ "Summer",
    Month %in% c(8, 9, 10) ~ "Fall",
    Month %in% c(11, 12, 1) ~ "Winter"
  ))
```

7.  Use `table` to figure out how many observations we have from each season.

```{r}
table(airquality$season)
```

There are 61 recorded obervations for Fall and 92 recorded observations for Summer, with 0 recorded for Spring and Winter. 

## **Part 3: Data Preprocessing**

8.  Normalize the predictor variables (Temp, Solar.R, Wind, and Season) using a `recipe`

```{r}
library(tidymodels)

(recipe_obj <- recipe(Ozone ~ Solar.R + Temp + Wind + season, data = airquality) |>
  step_impute_mean(all_numeric_predictors()) |>
  step_dummy(all_factor_predictors()) |>
  step_normalize(all_numeric_predictors())
  )
```

9.  What is the purpose of normalizing data?

Data normalization is important for creating models since it ensures that all of the variables are on a common scale, prevents biased modelling, and improves comparisons. 

10. What function can be used to impute missing values with the mean?

The step_impute_mean() fuction can be used to impute missing values with the mean. 

11. `prep` and `bake` the data to generate a processed dataset.

```{r}
prep_recipe <- prep(recipe_obj, training = airquality)

normalized_data <- bake(prep_recipe, new_data = NULL) |>
  drop_na()

normalized_data
```

12. Why is it necessary to both `prep()` and `bake()` the recipe?

It is necessary to prep and bake the recipe since that estimates the parameters for transformation and then applies the transformation to the data.

## **Part 4: Building a Linear Regression Model**

13. Fit a linear model using Ozone as the response variable and all other variables as predictors. Remember that the `.` notation can we used to include all variables.

```{r}
(model = lm(Ozone ~ ., data = normalized_data))

glance(model)

(pred <- augment(model, normalized_data))
```

14. Interpret the model summary output (coefficients, R-squared, p-values) in plain language

In the model summary output I received an r-squared value of 0.596, which demonstrates that 59.6% of the variability for Ozone can be attributed to the model. I also received a p-value of 4.58e-21 which indicated that the null hypothesis can be rejected with high confidence. The coefficients for the model include Intercept = 42.213, Solar.R = 4.857, Temp = 16.376, Wind = -10.952, and Season = 1.959.

## **Part 5: Model Diagnostics**

15. Use `broom::augment` to supplement the normalized `data.frame` with the fitted values and residuals.

```{r}
library(broom)

augmented_data <- augment(model, new_data = normalized_data)

augmented_data
```

16. Extract the residuals and visualize their distribution as a histogram and qqplot.

```{r}
residuals <- resid(model)

hist(residuals,
     main = "Residuals Histogram",
     xlab = "Residuals",
     col = "green",
     border = "black",
     breaks = 20)

qqnorm(residuals, main = "Residuals QQPlot")
qqline(residuals, col = "red")
```

17. Use `ggarange` to plot this as one image and interpret what you see in them.

```{r}
install.packages("ggpubr", repos = "https://cran.rstudio.com/")
library(ggpubr)

hist_plot <- ggplot(augmented_data, aes(x = residuals)) +
  geom_histogram(binwidth = 1, fill = "gray", color = "green") +
  labs(title = "Residuals Histogram", x = "Residuals", y = "Frequency") +
  theme_bw()

qq_plot <- ggplot(augmented_data, aes(sample = residuals)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Residuals QQPLOT") +
  theme_bw()

ggarrange(hist_plot, qq_plot,
          ncol = 2, nrow = 1,
          labels = c("A", "B"))
```

Based on these two graphs, the residuals seems to be normally distributed, though slightly skewed to the left, since they create a bell curve around 0 in the histogram graphs and follow the line fairly closely in the qqplot.  

18. Create a scatter plot of actual vs. predicted values using ggpubr with the following setting:

```{r}         
library(ggplot2)

ggscatter(augmented_data, x = "Ozone", y = ".fitted",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "spearman",
          ellipse = TRUE)
```

19. How strong of a model do you think this is?

I would say this is a fairly strong model because the plots on the graph are relatively clustered with few dramatic outliers and almost all of the plots are contained in the ellipse. Additionally, the R and p values indicate that the model strongly rejects the null hypothesis. 
