---
title: "daily_exercise_27"
format: html
editor: visual
---

## Library
```{r}
library(sf)   
library(osmdata)     
library(dplyr)    
library(terra)
library(ggplot2) 
library(viridis)
```

## Define River Object
```{r}
Sys.setenv("AWS_NO_SIGN_REQUEST" = "YES")

(bb = AOI::geocode("Fort Collins", bbox = TRUE) |> 
  st_transform(5070))

elev = rast('/vsis3/lynker-spatial/gridded-resources/dem.vrt') |> 
  crop(bb)

writeRaster(elev, filename = "data/foco-elev.tif", overwrite = TRUE)

writeRaster(elev, filename = "../resources/foco-elev-cm.tif", overwrite = TRUE)

fc = AOI::geocode("Fort Collins", bbox = TRUE) |> 
  st_transform(crs(elev))

foco_rivers <- osmdata::opq(st_bbox(st_transform(fc,4326))) |> 
  add_osm_feature("waterway") |> 
  osmdata_sf()

poudre_river <- foco_rivers$osm_lines |>
  filter(name == "Cache la Poudre River")

poudre_river_union <- st_union(poudre_river)

poudre_river_sf <- st_as_sf(poudre_river_union)

river_length <- st_length(poudre_river_sf)

poudre_pts <- st_cast(poudre_river_sf, "POINT")
```

## Define DEM Object
```{r}
elev <- rast("/vsis3/lynker-spatial/gridded-resources/dem.vrt")
```

## Extract River Profile
```{r}
poudre_pts <- st_transform(poudre_pts, crs(elev))

poudre_vect <- vect(poudre_pts)

elev_values <- terra::extract(elev, poudre_vect)

poudre_profile <- dplyr::bind_cols(poudre_pts, elev_values)

poudre_profile <- poudre_profile |> 
  mutate(ID = row_number())
```

## Compute Sinuosity
```{r}
first_point <- poudre_profile[1, ]  
last_point <- poudre_profile[nrow(poudre_profile), ]

straight_line_dist <- st_distance(first_point, last_point)

straight_line_dist <- as.numeric(st_distance(first_point, last_point))

river_length <- st_length(poudre_river_sf)

sinuosity <- as.numeric(river_length) / straight_line_dist
sinuosity
```

When computing the sinuosity of the Chache la Poudre River, I recieved a value of 2.897. This value is fairly typical for long, winding rivers and indicated that the river is 2.9 times longer than a straight line. Given that the Poudre runs through a canyon, this value aligns with the sinuosity of the river.

## Compute Slope
```{r}
first_elevation <- poudre_profile$dem[1]  
last_elevation <- poudre_profile$dem[nrow(poudre_profile)]

first_elevation <- first_elevation * 100  
last_elevation <- last_elevation * 100

elevation_change <- abs(last_elevation - first_elevation)

river_length_cm <- as.numeric(river_length) * 100

slope <- elevation_change / river_length_cm
slope
```

When computing the slope for the Cache la Poudre River, I recieved a value of 1.01 cm/m. This indicates that for every meter of river length, there is a 1.01 cm increase in elevation. This is fairly typical for flat, urban areas like Fort Collins. 

## Map Profile
```{r}
ggplot(poudre_profile, aes(x = ID, y = dem)) +
  geom_line(color = "blue", size = 1) +  
  geom_point(color = "red", size = 2) +  
  labs(
    title = "River Profile of Cache la Poudre River",
    x = "Point ID",
    y = "Elevation (m)",
    caption = "Elevation profile along the Cache la Poudre River"
  ) +
  theme_bw() +  
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) 
```
